### 2022-08-22 Bloods Data Sheet.Rmd
Imported Data

#Main Data Sheet
```{r}
main_df <- read.csv("C:\\Users\\thoma\\OneDrive - Murdoch University\\Desktop\\PhD 2021\\Study 2 and 3\\AAADATA\\Bloods Data\\MAIN_DATA_SHEET.csv")
```

#Calculating Dilution Factor
```{r}
main_df <- main_df%>%
  filter(!Sample %in% 22)%>%
  filter(!Sample %in% 235)%>%
  filter(!Sample %in% 236)%>%
  mutate(Dilution_Factor = ((Sample.Volume+Solution)/Sample.Volume))%>%
  mutate(cal_LDH = LDH*Dilution_Factor)%>%
  mutate(cal_CK = CK*Dilution_Factor)%>%
  mutate(cal_Na = Na*Dilution_Factor)%>%
  mutate(cal_K = K*Dilution_Factor)%>%
  mutate(cal_Cl = Cl*Dilution_Factor)%>%
  drop_na(Sample)

```

```{r}
histogram(main_df$cal_Na)
```

#Summary Data
```{r}
BlaSUM <- main_df%>%
summarySE(measurevar = "Bla", groupvar =c("Session"), na.rm = TRUE, conf.interval = 0.95, .drop = TRUE)

LDHSUM <- main_df%>%
summarySE(measurevar = "cal_LDH", groupvar =c("Session"), na.rm = TRUE, conf.interval = 0.95, .drop = TRUE)

CKSUM <- main_df%>%
summarySE(measurevar = "cal_CK", groupvar =c("Session"), na.rm = TRUE, conf.interval = 0.95, .drop = TRUE)

NaSUM <- main_df%>%
summarySE(measurevar = "cal_Na", groupvar =c("Session"), na.rm = TRUE, conf.interval = 0.95, .drop = TRUE)

KSUM <- main_df%>%
summarySE(measurevar = "cal_K", groupvar =c("Session"), na.rm = TRUE, conf.interval = 0.95, .drop = TRUE)

ClSUM <- main_df%>%
summarySE(measurevar = "cal_Cl", groupvar =c("Session"), na.rm = TRUE, conf.interval = 0.95, .drop = TRUE)
```