### 2022-08-23 Bloods Data Sheet.Rmd
data editing

#Main Data Sheet
```{r}
main_df <- read.csv("C:\\Users\\thoma\\OneDrive - Murdoch University\\Desktop\\PhD 2021\\Study 2 and 3\\AAADATA\\Bloods Data\\MAIN_DATA_SHEET.csv")
```

#Calculating Dilution Factor
```{r}
main_df <- main_df%>%
  #filter(!Sample %in% 22)%>%
  filter(!Sample %in% 235)%>%
  filter(!Sample %in% 236)%>%
  mutate(Dilution_Factor = ((Sample.Volume+Solution)/Sample.Volume))%>%
  mutate(cal_LDH = LDH*Dilution_Factor)%>%
  mutate(cal_CK = CK*Dilution_Factor)%>%
  mutate(cal_Na = (Na-0.68)*Dilution_Factor)%>%
  mutate(cal_K = (K-0.025)*Dilution_Factor)%>%
  mutate(cal_Cl = (Cl-2.8)*Dilution_Factor)%>%
  drop_na(Sample)
main_df$Time_Point <- as.factor(main_df$Time_Point)

```

```{r}
write.csv(main_df,"C:\\Users\\thoma\\OneDrive - Murdoch University\\Desktop\\PhD 2022\\Publications\\Study 2 - Bloods\\Main_Data_Frame (23-08-2022).csv",row.names = FALSE)
```

```{r}
histogram(main_df$cal_Na)
```

#Summary Data
```{r}
BlaSUM <- main_df%>%
summarySE(measurevar = "Bla", groupvar =c("Session"), na.rm = TRUE, conf.interval = 0.95, .drop = TRUE)

LDHSUM <- main_df%>%
summarySE(measurevar = "cal_LDH", groupvar =c("Session"), na.rm = TRUE, conf.interval = 0.95, .drop = TRUE)

CKSUM <- main_df%>%
summarySE(measurevar = "cal_CK", groupvar =c("Session"), na.rm = TRUE, conf.interval = 0.95, .drop = TRUE)

NaSUM <- main_df%>%
summarySE(measurevar = "cal_Na", groupvar =c("Session"), na.rm = TRUE, conf.interval = 0.95, .drop = TRUE)

KSUM <- main_df%>%
summarySE(measurevar = "cal_K", groupvar =c("Session"), na.rm = TRUE, conf.interval = 0.95, .drop = TRUE)

ClSUM <- main_df%>%
summarySE(measurevar = "cal_Cl", groupvar =c("Session"), na.rm = TRUE, conf.interval = 0.95, .drop = TRUE)
```

```{r}
main_df_PRE <- main_df%>%
  filter(!Time_Point %in% "2")%>%
  filter(!Time_Point %in% "3")

main_df <- main_df[main_df$cal_Cl <300,]
TRUNK_MAX_X <- TRUNK_MAX_X[TRUNK_MAX_X$TrunkAngle_X <10,]
```