### 2022-08-31 Bloods_LMM.Rmd

#Blood Lactate
```{r}
BlaLMM <- lmer(Bla ~ Time_Point + BFR + Speed + Time_Point*BFR + Time_Point*Speed + (1|Participant), data = main_df2)
summary(BlaLMM)
pairwise_Bla <- emmeans(BlaLMM, pairwise ~ Time_Point + BFR + Speed + Time_Point*BFR + Time_Point*Speed, adjust="holm")
pairwise_Bla
```

```{r}
tab_model(BlaLMM)
```

#Creatine Kinase
```{r}
CKLMM <- lmer(P_CHANGE ~ Time_Point + BFR + Speed + Time_Point*BFR + Time_Point*Speed + (1|Participant), data = CK_PCHANGE_CL)
summary(CKLMM)
pairwiseCK <- emmeans(CKLMM, pairwise ~ Time_Point + BFR + Speed + Time_Point*BFR + Time_Point*Speed, adjust="holm")
pairwiseCK
```
```{r}
tab_model(CKLMM)
```

#Lactate Dehydrogenase
```{r}
LDHLMM <- lmer(P_CHANGE ~ Time_Point + BFR + Speed + Time_Point*BFR + Time_Point*Speed + (1|Participant), data = LDH_PCHANGE_CL)
summary(LDHLMM)
pairwiseLDH <- emmeans(LDHLMM, pairwise ~ Time_Point + BFR + Speed + Time_Point*BFR + Time_Point*Speed, adjust="holm")
pairwiseLDH
```

```{r}
tab_model(LDHLMM)
```

```{r}
names(LDH_PCHANGE_CL)
```