### 2022-08-31 Draft Plots.Rmd

#Draft Plots
##Blood Lactate
```{r}
Bla_SUM<-main_df%>%
  summarySE(measurevar = "Bla", groupvar =c("Session","Time_Point", "Speed","BFR"), na.rm = TRUE, conf.interval = 0.95, .drop = TRUE)
```

```{r}
Bla_SUM$Session_f = factor(Bla_SUM$Session, levels=c("SW","SWBFR","FW", "FWBFR"))
 Session_f.labs<- c("Low-Intensity", "Low-Intensity + BFR", "Moderate-Intensity", "Moderate-Intensity + BFR")
 names(Session_f.labs)<-c("SW","SWBFR","FW", "FWBFR")

Bla_SUM$Speed_f = factor(Bla_SUM$Speed, levels=c("SLOW", "FAST"))
 Speed.labs<- c("Low-Intensity","Moderate-Intensity")
 names(Speed.labs)<-c("SLOW", "FAST")

Bla_SUM$Time_Point_f = factor(Bla_SUM$Time_Point, levels=c("1", "2", "3"))
Time_Point.labs<- c("Pre-Exercise", "Post-Exercise Finger", "Post-Exercise Toe")
names(Time_Point.labs)<-c("1", "2", "3")

FWBFR_Bla_Finger <- data.frame(Time_Point_f = c("2"),
                         Bla = c(6),
                         Session_f = c("FWBFR"))

FWBFR_Bla_Toe <- data.frame(Time_Point_f = c("3"),
                         Bla = c(6),
                         Session_f = c("FWBFR"))

FW_Bla_Toe <- data.frame(Time_Point_f = c("3"),
                         Bla = c(5),
                         Session_f = c("FW"))

SWBFR_Bla_Finger <- data.frame(Time_Point_f = c("2"),
                         Bla = c(4),
                         Session_f = c("SWBFR"))

SWBFR_Bla_Toe <- data.frame(Time_Point_f = c("3"),
                         Bla = c(4),
                         Session_f = c("SWBFR"))

Bla_PLOT_1 <- ggplot(Bla_SUM, aes(x = Time_Point_f, y= Bla))+
            geom_errorbar(aes(ymin = Bla-se, ymax =Bla+se, group = Session_f), width = 0.3, colour = "black", position = position_dodge(width = 0.5))+
            geom_point(aes(shape = Session_f), fill = "white", colour = "black", size = 5, stroke = 1.8, position = position_dodge(width = 0.5))+

            scale_colour_manual(name = "Session_f",
                            labels = Session_f.labs,
                            values = c("black", "black", "black", "black"))+

            scale_shape_manual(name = "Session_f",
                           labels = Session_f.labs,
                           values = c(21, 16,22, 15))+
            theme_classic()+
            theme(legend.position="bottom", legend.title=element_blank())+
            xlab("\n Collection Time Point")+
            scale_y_continuous(name = "Blood Lactate Concentration (mmol/L) \n",breaks = c(0,1,2,3,4,5,6), limits = c(0,6))+
            scale_x_discrete(breaks=c("1", "2", "3"),
              labels=c("Pre-Exercise", "Post-Exercise Finger", "Post-Exercise Toe"))+

                  theme(axis.line = element_line(colour = "black", size = 0.4, linetype = "solid"),
                  axis.text.x=element_text(size=16, colour="black"),
                  axis.text.y=element_text(size=16, colour="black"))+
                  theme(axis.text.x=element_text(size=16, angle = 0),
                   axis.title.x=element_text(size=16),
                   axis.title.y=element_text(size=18),
                   strip.text.y=element_text(size=14,face="italic"),
                   strip.text.x=element_text(size=16,face="italic"),
                   legend.title = element_blank())+
                   theme(legend.text=element_text(size=16))+

                  geom_text(size=6,data = FWBFR_Bla_Finger, label = "a #       ")+  
                   geom_text(size=6,data = FWBFR_Bla_Toe, label = "a  #     ")+
                   geom_text(size=6,data = FW_Bla_Toe, label = "a")+
                   geom_text(size=6,data = SWBFR_Bla_Finger, label = "#")+
                   geom_text(size=6,data = SWBFR_Bla_Toe, label = "#")

Bla_PLOT_1

```

```{r}
ggsave(plot=Bla_PLOT_1,"C://Users//thoma//OneDrive - Murdoch University//Desktop//PhD 2022//Publications//Study 2 - Bloods//Plot Drafts//Blaplot_Draft(31-08-2022).png", units = "in",width = 12, height = 8, dpi = 1200)

```

## Lactate Dehydrogenase
```{r}
LDH_SUM<-main_df%>%
  summarySE(measurevar = "cal_LDH", groupvar =c("Session","Time_Point", "Speed","BFR"), na.rm = TRUE, conf.interval = 0.95, .drop = TRUE)
```

```{r}
LDH_SUM$Session_f = factor(LDH_SUM$Session, levels=c("SW","SWBFR","FW", "FWBFR"))
 Session_f.labs<- c("Low-Intensity", "Low-Intensity + BFR", "Moderate-Intensity", "Moderate-Intensity + BFR")
 names(Session_f.labs)<-c("SW","SWBFR","FW", "FWBFR")

LDH_SUM$Speed_f = factor(LDH_SUM$Speed, levels=c("SW", "FW"))
 Speed.labs<- c("Low-Intensity","Moderate-Intensity")
 names(Speed.labs)<-c("SW", "FW")

LDH_SUM$Time_Point_f = factor(LDH_SUM$Time_Point, levels=c("1", "2", "3"))
Time_Point.labs<- c("Pre-Exercise", "Post-Exercise Finger", "Post-Exercise Toe")
names(Time_Point.labs)<-c("1", "2", "3")

LDH_PLOT_1 <- ggplot(LDH_SUM, aes(x = Time_Point_f, y= cal_LDH))+
            geom_errorbar(aes(ymin = cal_LDH-se, ymax =cal_LDH+se, group = Session), width = 0.3, colour = "black", position = position_dodge(width = 0.5))+
            geom_point(aes(shape = Session), fill = "white", colour = "black", size = 5, stroke = 1.8, position = position_dodge(width = 0.5))+

            scale_colour_manual(name = "Session_f",
                            labels = Session_f.labs,
                            values = c("black", "black", "black", "black"))+

            scale_shape_manual(name = "Session_f",
                           labels = Session_f.labs,
                           values = c(22, 15, 21, 16))+
            theme_classic()+
            theme(legend.position="right", legend.title=element_blank())+
            xlab("\n Collection Time Point")+
            scale_y_continuous(name = "Lactate Dehydrogenase (units/L) \n",breaks = c(150,160,170,180,190,200,210,220,230,240,250,260,270), limits = c(150,270))+
            scale_x_discrete(breaks=c("1", "2", "3"),
              labels=c("Pre-Exercise", "Post-Exercise Finger", "Post-Exercise Toe"))

LDH_PLOT_1

```

```{r}
ggsave(plot=LDH_PLOT_1,"C://Users//thoma//OneDrive - Murdoch University//Desktop//PhD 2022//Publications//Study 2 - Bloods//Plot Drafts//LDHplot_Draft(25-08-2022).png", units = "in",width = 12, height = 8, dpi = 1200)

```

### Lactate Dehydrogenase PLot 2
```{r}
LDH_SUM2<-main_df2%>%
  summarySE(measurevar = "cal_LDH", groupvar =c("Session","Time_Point", "Speed","BFR"), na.rm = TRUE, conf.interval = 0.95, .drop = TRUE)
```

```{r}
LDH_SUM2$Session_f = factor(LDH_SUM2$Session, levels=c("SW","SWBFR","FW", "FWBFR"))
 Session_f.labs<- c("Low-Intensity", "Low-Intensity + BFR", "Moderate-Intensity", "Moderate-Intensity + BFR")
 names(Session_f.labs)<-c("SW","SWBFR","FW", "FWBFR")

LDH_SUM2$Speed_f = factor(LDH_SUM2$Speed, levels=c("SW", "FW"))
 Speed.labs<- c("Low-Intensity","Moderate-Intensity")
 names(Speed.labs)<-c("SW", "FW")

LDH_SUM2$Time_Point_f = factor(LDH_SUM2$Time_Point, levels=c("1", "2", "3"))
Time_Point.labs<- c("Pre-Exercise", "Post-Exercise Finger", "Post-Exercise Toe")
names(Time_Point.labs)<-c("1", "2", "3")

LDH_PLOT_2 <- ggplot(LDH_SUM2, aes(x = Time_Point_f, y= cal_LDH))+
            geom_errorbar(aes(ymin = cal_LDH-se, ymax =cal_LDH+se, group = Session), width = 0.3, colour = "black", position = position_dodge(width = 0.5))+
            geom_point(aes(shape = Session), fill = "white", colour = "black", size = 5, stroke = 1.8, position = position_dodge(width = 0.5))+

            scale_colour_manual(name = "Session_f",
                            labels = Session_f.labs,
                            values = c("black", "black", "black", "black"))+

            scale_shape_manual(name = "Session_f",
                           labels = Session_f.labs,
                           values = c(22, 15, 21, 16))+
            theme_classic()+
            theme(legend.position="right", legend.title=element_blank())+
            xlab("\n Collection Time Point")+
            scale_y_continuous(name = "Lactate Dehydrogenase (units/L) \n",breaks = c(150,160,170,180,190,200,210,220,230,240,250,260,270), limits = c(150,270))+
            scale_x_discrete(breaks=c("1", "2", "3"),
              labels=c("Pre-Exercise", "Post-Exercise Finger", "Post-Exercise Toe"))

LDH_PLOT_2

```
```{r}
LDH_Grouping <- ggarrange(LDH_PLOT_1,LDH_PLOT_2,labels=c("A", "B"), ncol=2, nrow=1, common.legend = TRUE, legend = "bottom")

LDH_Grouping
```

```{r}
ggsave(plot=LDH_Grouping,"C://Users//thoma//OneDrive - Murdoch University//Desktop//PhD 2022//Publications//Study 2 - Bloods//Plot Drafts//LDHgrouping_Draft(26-08-2022).png", units = "in",width = 12, height = 8, dpi = 1200)

```

##Creatine Kinase
```{r}
CK_SUM<-main_df%>%
  summarySE(measurevar = "cal_CK", groupvar =c("Session","Time_Point", "Speed","BFR"), na.rm = TRUE, conf.interval = 0.95, .drop = TRUE)
```

```{r}
CK_SUM$Session_f = factor(CK_SUM$Session, levels=c("SW","SWBFR","FW", "FWBFR"))
 Session_f.labs<- c("Low-Intensity", "Low-Intensity + BFR", "Moderate-Intensity", "Moderate-Intensity + BFR")
 names(Session_f.labs)<-c("SW","SWBFR","FW", "FWBFR")

CK_SUM$Speed_f = factor(CK_SUM$Speed, levels=c("SW", "FW"))
 Speed.labs<- c("Low-Intensity","Moderate-Intensity")
 names(Speed.labs)<-c("SW", "FW")

CK_SUM$Time_Point_f = factor(CK_SUM$Time_Point, levels=c("1", "2", "3"))
Time_Point.labs<- c("Pre-Exercise", "Post-Exercise Finger", "Post-Exercise Toe")
names(Time_Point.labs)<-c("1", "2", "3")

CK_PLOT_1 <- ggplot(CK_SUM, aes(x = Time_Point_f, y= cal_CK))+
            geom_errorbar(aes(ymin = cal_CK-se, ymax =cal_CK+se, group = Session), width = 0.3, colour = "black", position = position_dodge(width = 0.5))+
            geom_point(aes(shape = Session), fill = "white", colour = "black", size = 5, stroke = 1.8, position = position_dodge(width = 0.5))+

            scale_colour_manual(name = "Session_f",
                            labels = Session_f.labs,
                            values = c("black", "black", "black", "black"))+

            scale_shape_manual(name = "Session_f",
                           labels = Session_f.labs,
                           values = c(22, 15, 21, 16))+
            theme_classic()+
            theme(legend.position="right", legend.title=element_blank())+
            xlab("\n Collection Time Point")+
            scale_y_continuous(name = "Creatine Kinase (units/L) \n",breaks = c(125,150,175,200,225,250,275,300,325,350), limits = c(125,350))+
            scale_x_discrete(breaks=c("1", "2", "3"),
              labels=c("Pre-Exercise", "Post-Exercise Finger", "Post-Exercise Toe"))

CK_PLOT_1

```

```{r}
ggsave(plot=CK_PLOT_1,"C://Users//thoma//OneDrive - Murdoch University//Desktop//PhD 2022//Publications//Study 2 - Bloods//Plot Drafts//CKplot_Draft(25-08-2022).png", units = "in",width = 12, height = 8, dpi = 1200)

```

###CK_PLOT2
```{r}
CK_SUM2<-main_df2%>%
  summarySE(measurevar = "cal_CK", groupvar =c("Session","Time_Point", "Speed","BFR"), na.rm = TRUE, conf.interval = 0.95, .drop = TRUE)
```

```{r}
CK_SUM2$Session_f = factor(CK_SUM2$Session, levels=c("SW","SWBFR","FW", "FWBFR"))
 Session_f.labs<- c("Low-Intensity", "Low-Intensity + BFR", "Moderate-Intensity", "Moderate-Intensity + BFR")
 names(Session_f.labs)<-c("SW","SWBFR","FW", "FWBFR")

CK_SUM2$Speed_f = factor(CK_SUM2$Speed, levels=c("SW", "FW"))
 Speed.labs<- c("Low-Intensity","Moderate-Intensity")
 names(Speed.labs)<-c("SW", "FW")

CK_SUM2$Time_Point_f = factor(CK_SUM2$Time_Point, levels=c("1", "2", "3"))
Time_Point.labs<- c("Pre-Exercise", "Post-Exercise Finger", "Post-Exercise Toe")
names(Time_Point.labs)<-c("1", "2", "3")

CK_PLOT_2 <- ggplot(CK_SUM2, aes(x = Time_Point_f, y= cal_CK))+
            geom_errorbar(aes(ymin = cal_CK-se, ymax =cal_CK+se, group = Session), width = 0.3, colour = "black", position = position_dodge(width = 0.5))+
            geom_point(aes(shape = Session), fill = "white", colour = "black", size = 5, stroke = 1.8, position = position_dodge(width = 0.5))+

            scale_colour_manual(name = "Session_f",
                            labels = Session_f.labs,
                            values = c("black", "black", "black", "black"))+

            scale_shape_manual(name = "Session_f",
                           labels = Session_f.labs,
                           values = c(22, 15, 21, 16))+
            theme_classic()+
            theme(legend.position="right", legend.title=element_blank())+
            xlab("\n Collection Time Point")+
            scale_y_continuous(name = "Creatine Kinase (units/L) \n",breaks = c(100,125,150,175,200,225,250,275,300,325,350), limits = c(100,350))+
            scale_x_discrete(breaks=c("1", "2", "3"),
              labels=c("Pre-Exercise", "Post-Exercise Finger", "Post-Exercise Toe"))

CK_PLOT_2

```

```{r}
ggsave(plot=CK_PLOT_1,"C://Users//thoma//OneDrive - Murdoch University//Desktop//PhD 2022//Publications//Study 2 - Bloods//Plot Drafts//CKplot_Draft(25-08-2022).png", units = "in",width = 12, height = 8, dpi = 1200)

```

```{r}
CK_Grouping <- ggarrange(CK_PLOT_1,CK_PLOT_2,labels=c("A", "B"), ncol=2, nrow=1, common.legend = TRUE, legend = "bottom")

CK_Grouping
```
```{r}
ggsave(plot=CK_Grouping,"C://Users//thoma//OneDrive - Murdoch University//Desktop//PhD 2022//Publications//Study 2 - Bloods//Plot Drafts//CKgrouping_Draft(26-08-2022).png", units = "in",width = 12, height = 8, dpi = 1200)

```

##Sodium
```{r}
Na_SUM<-main_df%>%
  summarySE(measurevar = "cal_Na", groupvar =c("Session","Time_Point", "Speed","BFR"), na.rm = TRUE, conf.interval = 0.95, .drop = TRUE)
```

```{r}
Na_SUM$Session_f = factor(Na_SUM$Session, levels=c("SW","SWBFR","FW", "FWBFR"))
 Session_f.labs<- c("Low-Intensity", "Low-Intensity + BFR", "Moderate-Intensity", "Moderate-Intensity + BFR")
 names(Session_f.labs)<-c("SW","SWBFR","FW", "FWBFR")

Na_SUM$Speed_f = factor(Na_SUM$Speed, levels=c("SW", "FW"))
 Speed.labs<- c("Low-Intensity","Moderate-Intensity")
 names(Speed.labs)<-c("SW", "FW")

Na_SUM$Time_Point_f = factor(Na_SUM$Time_Point, levels=c("1", "2", "3"))
Time_Point.labs<- c("Pre-Exercise", "Post-Exercise Finger", "Post-Exercise Toe")
names(Time_Point.labs)<-c("1", "2", "3")

Na_PLOT_1 <- ggplot(Na_SUM, aes(x = Time_Point_f, y= cal_Na))+
            geom_errorbar(aes(ymin = cal_Na-se, ymax =cal_Na+se, group = Session), width = 0.3, colour = "black", position = position_dodge(width = 0.5))+
            geom_point(aes(shape = Session), fill = "white", colour = "black", size = 5, stroke = 1.8, position = position_dodge(width = 0.5))+

            scale_colour_manual(name = "Session_f",
                            labels = Session_f.labs,
                            values = c("black", "black", "black", "black"))+

            scale_shape_manual(name = "Session_f",
                           labels = Session_f.labs,
                           values = c(22, 15, 21, 16))+
            theme_classic()+
            theme(legend.position="right", legend.title=element_blank())+
            xlab("\n Collection Time Point")+
            scale_y_continuous(name = "Sodium \n",breaks = c(125,150,175,200), limits = c(125,200))+
            scale_x_discrete(breaks=c("1", "2", "3"),
              labels=c("Pre-Exercise", "Post-Exercise Finger", "Post-Exercise Toe"))

Na_PLOT_1

```

```{r}
ggsave(plot=Na_PLOT_1,"C://Users//thoma//OneDrive - Murdoch University//Desktop//PhD 2022//Publications//Study 2 - Bloods//Plot Drafts//NAplot_Draft(25-08-2022).png", units = "in",width = 12, height = 8, dpi = 1200)

```

###Sodum PLot 2
```{r}
Na_SUM2<-main_df2%>%
  summarySE(measurevar = "cal_Na", groupvar =c("Session","Time_Point", "Speed","BFR"), na.rm = TRUE, conf.interval = 0.95, .drop = TRUE)
```

```{r}
Na_SUM2$Session_f = factor(Na_SUM2$Session, levels=c("SW","SWBFR","FW", "FWBFR"))
 Session_f.labs<- c("Low-Intensity", "Low-Intensity + BFR", "Moderate-Intensity", "Moderate-Intensity + BFR")
 names(Session_f.labs)<-c("SW","SWBFR","FW", "FWBFR")

Na_SUM2$Speed_f = factor(Na_SUM2$Speed, levels=c("SW", "FW"))
 Speed.labs<- c("Low-Intensity","Moderate-Intensity")
 names(Speed.labs)<-c("SW", "FW")

Na_SUM2$Time_Point_f = factor(Na_SUM2$Time_Point, levels=c("1", "2", "3"))
Time_Point.labs<- c("Pre-Exercise", "Post-Exercise Finger", "Post-Exercise Toe")
names(Time_Point.labs)<-c("1", "2", "3")

Na_PLOT_2 <- ggplot(Na_SUM2, aes(x = Time_Point_f, y= cal_Na))+
            geom_errorbar(aes(ymin = cal_Na-se, ymax =cal_Na+se, group = Session_f), width = 0.3, colour = "black", position = position_dodge(width = 0.5))+
            geom_point(aes(shape = Session_f), fill = "white", colour = "black", size = 5, stroke = 1.8, position = position_dodge(width = 0.5))+

            scale_colour_manual(name = "Session_f",
                            labels = Session_f.labs,
                            values = c("black", "black", "black", "black"))+

            scale_shape_manual(name = "Session_f",
                           labels = Session_f.labs,
                           values = c(21, 16, 22, 15))+
            theme_classic()+
            theme(legend.position="", legend.title=element_blank())+
            xlab("\n Collection Time Point")+
            scale_y_continuous(name = "Sodium (mmol/L) \n",breaks = c(100,125,150,175,200), limits = c(100,200))+
            scale_x_discrete(breaks=c("1", "2", "3"),
              labels=c("Pre-Exercise", "Post-Exercise Finger", "Post-Exercise Toe"))+

  theme(axis.line = element_line(colour = "black", size = 0.4, linetype = "solid"),
                  axis.text.x=element_text(size=16, colour="black"),
                  axis.text.y=element_text(size=16, colour="black"))+
                  theme(axis.text.x=element_text(size=10, angle = 0),
                   axis.title.x=element_text(size=16),
                   axis.title.y=element_text(size=18),
                   strip.text.y=element_text(size=14,face="italic"),
                   strip.text.x=element_text(size=16,face="italic"),
                   legend.title = element_blank())+
                   theme(legend.text=element_text(size=16))

Na_PLOT_2

```

```{r}
ggsave(plot=Na_PLOT_1,"C://Users//thoma//OneDrive - Murdoch University//Desktop//PhD 2022//Publications//Study 2 - Bloods//Plot Drafts//NAplot_Draft(25-08-2022).png", units = "in",width = 12, height = 8, dpi = 1200)

```

```{r}
NA_Grouping <- ggarrange(Na_PLOT_1,Na_PLOT_2,labels=c("A", "B"), ncol=2, nrow=1, common.legend = TRUE, legend = "bottom")

NA_Grouping
```

```{r}
ggsave(plot=NA_Grouping,"C://Users//thoma//OneDrive - Murdoch University//Desktop//PhD 2022//Publications//Study 2 - Bloods//Plot Drafts//Nagrouping_Draft(26-08-2022).png", units = "in",width = 12, height = 8, dpi = 1200)

```

##Potassium
```{r}
K_SUM<-main_df%>%
  summarySE(measurevar = "cal_K", groupvar =c("Session","Time_Point", "Speed","BFR"), na.rm = TRUE, conf.interval = 0.95, .drop = TRUE)
```

```{r}
K_SUM$Session_f = factor(K_SUM$Session, levels=c("SW","SWBFR","FW", "FWBFR"))
 Session_f.labs<- c("Low-Intensity", "Low-Intensity + BFR", "Moderate-Intensity", "Moderate-Intensity + BFR")
 names(Session_f.labs)<-c("SW","SWBFR","FW", "FWBFR")

K_SUM$Speed_f = factor(K_SUM$Speed, levels=c("SW", "FW"))
 Speed.labs<- c("Low-Intensity","Moderate-Intensity")
 names(Speed.labs)<-c("SW", "FW")

K_SUM$Time_Point_f = factor(K_SUM$Time_Point, levels=c("1", "2", "3"))
Time_Point.labs<- c("Pre-Exercise", "Post-Exercise Finger", "Post-Exercise Toe")
names(Time_Point.labs)<-c("1", "2", "3")

K_PLOT_1 <- ggplot(K_SUM, aes(x = Time_Point_f, y= cal_K))+
            geom_errorbar(aes(ymin = cal_K-se, ymax =cal_K+se, group = Session), width = 0.3, colour = "black", position = position_dodge(width = 0.5))+
            geom_point(aes(shape = Session), fill = "white", colour = "black", size = 5, stroke = 1.8, position = position_dodge(width = 0.5))+

            scale_colour_manual(name = "Session_f",
                            labels = Session_f.labs,
                            values = c("black", "black", "black", "black"))+

            scale_shape_manual(name = "Session_f",
                           labels = Session_f.labs,
                           values = c(22, 15, 21, 16))+
            theme_classic()+
            theme(legend.position="right", legend.title=element_blank())+
            xlab("\n Collection Time Point")+
            scale_y_continuous(name = "Potassium \n",breaks = c(4,5,6,7,8,9,10), limits = c(4,10))+
            scale_x_discrete(breaks=c("1", "2", "3"),
              labels=c("Pre-Exercise", "Post-Exercise Finger", "Post-Exercise Toe"))

K_PLOT_1

```

```{r}
ggsave(plot=K_PLOT_1,"C://Users//thoma//OneDrive - Murdoch University//Desktop//PhD 2022//Publications//Study 2 - Bloods//Plot Drafts//Kplot_Draft(25-08-2022).png", units = "in",width = 12, height = 8, dpi = 1200)

```

###Potassium Plot 2
```{r}
K_SUM2<-main_df2%>%
  summarySE(measurevar = "cal_K", groupvar =c("Session","Time_Point", "Speed","BFR"), na.rm = TRUE, conf.interval = 0.95, .drop = TRUE)
```

```{r}
K_SUM2$Session_f = factor(K_SUM2$Session, levels=c("SW","SWBFR","FW", "FWBFR"))
 Session_f.labs<- c("Low-Intensity", "Low-Intensity + BFR", "Moderate-Intensity", "Moderate-Intensity + BFR")
 names(Session_f.labs)<-c("SW","SWBFR","FW", "FWBFR")

K_SUM2$Speed_f = factor(K_SUM2$Speed, levels=c("SW", "FW"))
 Speed.labs<- c("Low-Intensity","Moderate-Intensity")
 names(Speed.labs)<-c("SW", "FW")

K_SUM2$Time_Point_f = factor(K_SUM2$Time_Point, levels=c("1", "2", "3"))
Time_Point.labs<- c("Pre-Exercise", "Post-Exercise Finger", "Post-Exercise Toe")
names(Time_Point.labs)<-c("1", "2", "3")

K_PLOT_2 <- ggplot(K_SUM2, aes(x = Time_Point_f, y= cal_K))+
            geom_errorbar(aes(ymin = cal_K-se, ymax =cal_K+se, group = Session_f), width = 0.3, colour = "black", position = position_dodge(width = 0.5))+
            geom_point(aes(shape = Session_f), fill = "white", colour = "black", size = 5, stroke = 1.8, position = position_dodge(width = 0.5))+

            scale_colour_manual(name = "Session_f",
                            labels = Session_f.labs,
                            values = c("black", "black", "black", "black"))+

            scale_shape_manual(name = "Session_f",
                           labels = Session_f.labs,
                           values = c(21, 16, 22, 15))+
            theme_classic()+
            theme(legend.position="", legend.title=element_blank())+
            xlab("\n Collection Time Point")+
            scale_y_continuous(name = "Potassium (mmol/L) \n",breaks = c(4,5,6,7,8,9,10), limits = c(4,10))+
            scale_x_discrete(breaks=c("1", "2", "3"),
              labels=c("Pre-Exercise", "Post-Exercise Finger", "Post-Exercise Toe"))+

  theme(axis.line = element_line(colour = "black", size = 0.4, linetype = "solid"),
                  axis.text.x=element_text(size=16, colour="black"),
                  axis.text.y=element_text(size=16, colour="black"))+
                  theme(axis.text.x=element_text(size=10, angle = 0),
                   axis.title.x=element_text(size=16),
                   axis.title.y=element_text(size=18),
                   strip.text.y=element_text(size=14,face="italic"),
                   strip.text.x=element_text(size=16,face="italic"),
                   legend.title = element_blank())+
                   theme(legend.text=element_text(size=16))

K_PLOT_2

```

```{r}
ggsave(plot=K_PLOT_1,"C://Users//thoma//OneDrive - Murdoch University//Desktop//PhD 2022//Publications//Study 2 - Bloods//Plot Drafts//Kplot_Draft(25-08-2022).png", units = "in",width = 12, height = 8, dpi = 1200)

```

```{r}
K_Grouping <- ggarrange(K_PLOT_1,K_PLOT_2,labels=c("A", "B"), ncol=2, nrow=1, common.legend = TRUE, legend = "bottom")

K_Grouping
```

```{r}
ggsave(plot=K_Grouping,"C://Users//thoma//OneDrive - Murdoch University//Desktop//PhD 2022//Publications//Study 2 - Bloods//Plot Drafts//Kgrouping_Draft(26-08-2022).png", units = "in",width = 12, height = 8, dpi = 1200)

```

##Chlorine
```{r}
Cl_SUM<-main_df%>%
  summarySE(measurevar = "cal_Cl", groupvar =c("Session","Time_Point", "Speed","BFR"), na.rm = TRUE, conf.interval = 0.95, .drop = TRUE)
```

```{r}
Cl_SUM$Session_f = factor(Cl_SUM$Session, levels=c("SW","SWBFR","FW", "FWBFR"))
 Session_f.labs<- c("Low-Intensity", "Low-Intensity + BFR", "Moderate-Intensity", "Moderate-Intensity + BFR")
 names(Session_f.labs)<-c("SW","SWBFR","FW", "FWBFR")

Cl_SUM$Speed_f = factor(Cl_SUM$Speed, levels=c("SW", "FW"))
 Speed.labs<- c("Low-Intensity","Moderate-Intensity")
 names(Speed.labs)<-c("SW", "FW")

Cl_SUM$Time_Point_f = factor(Cl_SUM$Time_Point, levels=c("1", "2", "3"))
Time_Point.labs<- c("Pre-Exercise", "Post-Exercise Finger", "Post-Exercise Toe")
names(Time_Point.labs)<-c("1", "2", "3")

Cl_PLOT_1 <- ggplot(Cl_SUM, aes(x = Time_Point_f, y= cal_Cl))+
            geom_errorbar(aes(ymin = cal_Cl-se, ymax =cal_Cl+se, group = Session), width = 0.3, colour = "black", position = position_dodge(width = 0.5))+
            geom_point(aes(shape = Session), fill = "white", colour = "black", size = 5, stroke = 1.8, position = position_dodge(width = 0.5))+

            scale_colour_manual(name = "Session_f",
                            labels = Session_f.labs,
                            values = c("black", "black", "black", "black"))+

            scale_shape_manual(name = "Session_f",
                           labels = Session_f.labs,
                           values = c(22, 15, 21, 16))+
            theme_classic()+
            theme(legend.position="right", legend.title=element_blank())+
            xlab("\n Collection Time Point")+
            scale_y_continuous(name = "Chlorine \n",breaks = c(90,100,110,120), limits = c(90,120))+
            scale_x_discrete(breaks=c("1", "2", "3"),
              labels=c("Pre-Exercise", "Post-Exercise Finger", "Post-Exercise Toe"))

Cl_PLOT_1

```

```{r}
ggsave(plot=Cl_PLOT_1,"C://Users//thoma//OneDrive - Murdoch University//Desktop//PhD 2022//Publications//Study 2 - Bloods//Plot Drafts//Clplot_Draft(25-08-2022).png", units = "in",width = 12, height = 8, dpi = 1200)

```

###Chorine PLot 2
```{r}
Cl_SUM2<-main_df2%>%
  summarySE(measurevar = "cal_Cl", groupvar =c("Session","Time_Point", "Speed","BFR"), na.rm = TRUE, conf.interval = 0.95, .drop = TRUE)
```

```{r}
Cl_SUM2$Session_f = factor(Cl_SUM2$Session, levels=c("SW","SWBFR","FW", "FWBFR"))
 Session_f.labs<- c("Low-Intensity", "Low-Intensity + BFR", "Moderate-Intensity", "Moderate-Intensity + BFR")
 names(Session_f.labs)<-c("SW","SWBFR","FW", "FWBFR")

Cl_SUM2$Speed_f = factor(Cl_SUM2$Speed, levels=c("SW", "FW"))
 Speed.labs<- c("Low-Intensity","Moderate-Intensity")
 names(Speed.labs)<-c("SW", "FW")

Cl_SUM2$Time_Point_f = factor(Cl_SUM2$Time_Point, levels=c("1", "2", "3"))
Time_Point.labs<- c("Pre-Exercise", "Post-Exercise Finger", "Post-Exercise Toe")
names(Time_Point.labs)<-c("1", "2", "3")

Cl_PLOT_2 <- ggplot(Cl_SUM2, aes(x = Time_Point_f, y= cal_Cl))+
            geom_errorbar(aes(ymin = cal_Cl-se, ymax =cal_Cl+se, group = Session_f), width = 0.3, colour = "black", position = position_dodge(width = 0.5))+
            geom_point(aes(shape = Session_f), fill = "white", colour = "black", size = 5, stroke = 1.8, position = position_dodge(width = 0.5))+

            scale_colour_manual(name = "Session_f",
                            labels = Session_f.labs,
                            values = c("black", "black", "black", "black"))+

            scale_shape_manual(name = "Session_f",
                           labels = Session_f.labs,
                           values = c(21, 16, 22, 15))+
            theme_classic()+
            theme(legend.position="", legend.title=element_blank())+
            xlab("\n Collection Time Point")+
            scale_y_continuous(name = "Chlorine (mmol/L) \n",breaks = c(50,75,100,125,150), limits = c(50,150))+
            scale_x_discrete(breaks=c("1", "2", "3"),
              labels=c("Pre-Exercise", "Post-Exercise Finger", "Post-Exercise Toe"))+
  theme(axis.line = element_line(colour = "black", size = 0.4, linetype = "solid"),
                  axis.text.x=element_text(size=16, colour="black"),
                  axis.text.y=element_text(size=16, colour="black"))+
                  theme(axis.text.x=element_text(size=10, angle = 0),
                   axis.title.x=element_text(size=16),
                   axis.title.y=element_text(size=18),
                   strip.text.y=element_text(size=14,face="italic"),
                   strip.text.x=element_text(size=16,face="italic"),
                   legend.title = element_blank())+
                   theme(legend.text=element_text(size=16))

Cl_PLOT_2

```

```{r}
ggsave(plot=Cl_PLOT_1,"C://Users//thoma//OneDrive - Murdoch University//Desktop//PhD 2022//Publications//Study 2 - Bloods//Plot Drafts//Clplot_Draft(25-08-2022).png", units = "in",width = 12, height = 8, dpi = 1200)

```

```{r}
Cl_Grouping <- ggarrange(Cl_PLOT_1,Cl_PLOT_2,labels=c("A", "B"), ncol=2, nrow=1, common.legend = TRUE, legend = "bottom")

Cl_Grouping
```

```{r}
ggsave(plot=Cl_Grouping,"C://Users//thoma//OneDrive - Murdoch University//Desktop//PhD 2022//Publications//Study 2 - Bloods//Plot Drafts//Clgrouping_Draft(26-08-2022).png", units = "in",width = 12, height = 8, dpi = 1200)

```

#Percentage Change
##CK
CK_PCHANGET2_SUM

```{r}
CK_PCHANGET2_SUM$Session_f = factor(CK_PCHANGET2_SUM$Session, levels=c("SW","SWBFR","FW", "FWBFR"))
 Session_f.labs<- c("Low-Intensity", "Low-Intensity + BFR", "Moderate-Intensity", "Moderate-Intensity + BFR")
 names(Session_f.labs)<-c("SW","SWBFR","FW", "FWBFR")

#Cl_SUM2$Speed_f = factor(Cl_SUM2$Speed, levels=c("SW", "FW"))
 #Speed.labs<- c("Low-Intensity","Moderate-Intensity")
 #names(Speed.labs)<-c("SW", "FW")

CK_PCHANGET2_SUM$Time_Point_f = factor(CK_PCHANGET2_SUM$Time_Point, levels=c("2", "3"))
Time_Point.labs<- c("Post-Exercise Finger", "Post-Exercise Toe")
names(Time_Point.labs)<-c("2", "3")

CK_PLOT_PCHANGE <- ggplot(CK_PCHANGET2_SUM, aes(x = Time_Point_f, y= P_CHANGE))+
            geom_errorbar(aes(ymin = P_CHANGE-se, ymax =P_CHANGE+se, group = Session_f), width = 0.3, colour = "black", position = position_dodge(width = 0.5))+
            geom_point(aes(shape = Session_f), fill = "white", colour = "black", size = 5, stroke = 1.8, position = position_dodge(width = 0.5))+

            scale_colour_manual(name = "Session_f",
                            labels = Session_f.labs,
                            values = c("black", "black", "black", "black"))+

            scale_shape_manual(name = "Session_f",
                           labels = Session_f.labs,
                           values = c(21, 16, 22, 15))+
            theme_classic()+
            ggtitle("Creatine Kinase")+
            theme(plot.title = element_text(hjust = 0.5))+
            theme(legend.position="", legend.title=element_blank())+
            xlab("\n Collection Time Point")+
            scale_y_continuous(name = "Percentage Change \n",breaks = c(-10,0,10,20,30,40,50), limits = c(-10,50))+
            scale_x_discrete(breaks=c("2", "3"),
              labels=c("Post-Exercise Finger", "Post-Exercise Toe"))+

             theme(axis.line = element_line(colour = "black", size = 0.4, linetype = "solid"),
                  axis.text.x=element_text(size=16, colour="black"),
                  axis.text.y=element_text(size=16, colour="black"))+
                  theme(axis.text.x=element_text(size=16, angle = 0),
                   axis.title.x=element_text(size=16),
                   axis.title.y=element_text(size=18),
                   strip.text.y=element_text(size=14,face="italic"),
                   strip.text.x=element_text(size=16,face="italic"),
                   legend.title = element_blank())+
                   theme(legend.text=element_text(size=16))

CK_PLOT_PCHANGE

```
##LDH
```{r}
LDH_PCHANGET2_SUM
```

```{r}
LDH_PCHANGET2_SUM$Session_f = factor(LDH_PCHANGET2_SUM$Session, levels=c("SW","SWBFR","FW", "FWBFR"))
 Session_f.labs<- c("Low-Intensity", "Low-Intensity + BFR", "Moderate-Intensity", "Moderate-Intensity + BFR")
 names(Session_f.labs)<-c("SW","SWBFR","FW", "FWBFR")

#Cl_SUM2$Speed_f = factor(Cl_SUM2$Speed, levels=c("SW", "FW"))
 #Speed.labs<- c("Low-Intensity","Moderate-Intensity")
 #names(Speed.labs)<-c("SW", "FW")

LDH_PCHANGET2_SUM$Time_Point_f = factor(LDH_PCHANGET2_SUM$Time_Point, levels=c("2", "3"))
Time_Point.labs<- c("Post-Exercise Finger", "Post-Exercise Toe")
names(Time_Point.labs)<-c("2", "3")

LDH_PLOT_PCHANGE <- ggplot(LDH_PCHANGET2_SUM, aes(x = Time_Point_f, y= P_CHANGE))+
            geom_errorbar(aes(ymin = P_CHANGE-se, ymax =P_CHANGE+se, group = Session_f), width = 0.3, colour = "black", position = position_dodge(width = 0.5))+
            geom_point(aes(shape = Session_f), fill = "white", colour = "black", size = 5, stroke = 1.8, position = position_dodge(width = 0.5))+

            scale_colour_manual(name = "Session_f",
                            labels = Session_f.labs,
                            values = c("black", "black", "black", "black"))+

            scale_shape_manual(name = "Session_f",
                           labels = Session_f.labs,
                           values = c(21, 16,22, 15))+
            theme_classic()+
            ggtitle("Lactate Dehydrogenase")+
            theme(plot.title = element_text(hjust = 0.5))+
            theme(legend.position="", legend.title=element_blank())+
            xlab("\n Collection Time Point")+
            scale_y_continuous(name = "Percentage Change \n",breaks = c(-10,0,10,20,30,40,50), limits = c(-10,50))+
            scale_x_discrete(breaks=c("2", "3"),
              labels=c("Post-Exercise Finger", "Post-Exercise Toe"))+

            theme(axis.line = element_line(colour = "black", size = 0.4, linetype = "solid"),
                  axis.text.x=element_text(size=16, colour="black"),
                  axis.text.y=element_text(size=16, colour="black"))+
                  theme(axis.text.x=element_text(size=16, angle = 0),
                   axis.title.x=element_text(size=16),
                   axis.title.y=element_text(size=18),
                   strip.text.y=element_text(size=14,face="italic"),
                   strip.text.x=element_text(size=16,face="italic"),
                   legend.title = element_blank())+
                   theme(legend.text=element_text(size=16))

LDH_PLOT_PCHANGE

```

```{r}
PCHANGE_Grouping <- ggarrange(CK_PLOT_PCHANGE,LDH_PLOT_PCHANGE,labels=c("A", "B"), ncol=2, nrow=1, common.legend = TRUE, legend = "bottom")

PCHANGE_Grouping
```

```{r}
ggsave(plot=PCHANGE_Grouping,"C://Users//thoma//OneDrive - Murdoch University//Desktop//PhD 2022//Publications//Study 2 - Bloods//Plot Drafts//PCHANGE_Grouping_Draft(31-08-2022).png", units = "in",width = 12, height = 8, dpi = 1200)

```

##Electrolyte Grouping
```{r}
ELECTROLYTE_Grouping <- ggarrange(Na_PLOT_2,Cl_PLOT_2,K_PLOT_2,labels=c("A", "B", "C"), ncol=3, nrow=1, common.legend = TRUE, legend = "bottom")

ELECTROLYTE_Grouping
```

```{r}
ggsave(plot=ELECTROLYTE_Grouping,"C://Users//thoma//OneDrive - Murdoch University//Desktop//PhD 2022//Publications//Study 2 - Bloods//Plot Drafts//ELECTROLYTE_Grouping(31-08-2022).png", units = "in",width = 18, height = 8, dpi = 1200)

```