### 2022-09-05 Bloods_LMM.Rmd
tried different model

#Blood Lactate
```{r}
BlaLMM <- lmer(Bla ~ Time_Point + BFR + Speed + Time_Point*BFR + Time_Point*Speed + (1|Participant), data = main_df2)
summary(BlaLMM)
pairwise_Bla <- emmeans(BlaLMM, pairwise ~ Time_Point + BFR + Speed + Time_Point*BFR + Time_Point*Speed, adjust="holm")
pairwise_Bla
```

```{r}
tab_model(BlaLMM)
```

#Creatine Kinase - Percentage Change
```{r}
CKLMM <- lmer(P_CHANGE ~ Time_Point + BFR + Speed + Time_Point*BFR + Time_Point*Speed + (1|Participant), data = CK_PCHANGE_CL)
summary(CKLMM)
pairwiseCK <- emmeans(CKLMM, pairwise ~ Time_Point + BFR + Speed + Time_Point*BFR + Time_Point*Speed, adjust="holm")
pairwiseCK
```
```{r}
tab_model(CKLMM)
```

#Creatine Kinase - Total Change
```{r}
#Cant use total change as the SWBFR is uncharacteristically low
```

```{r}
main_df_noPRE<-main_df2%>%
  filter(!Time_Point %in% "1")
```

```{r}
CKLMM_Total <- lmer(cal_CK ~ Time_Point + BFR + Speed + Time_Point*BFR + Time_Point*Speed + (1|Participant), data = main_df_noPRE)
summary(CKLMM_Total)
pairwiseCK_total <- emmeans(CKLMM_Total, pairwise ~ Time_Point + BFR + Speed + Time_Point*BFR + Time_Point*Speed, adjust="holm")
pairwiseCK_total
```
```{r}
tab_model(CKLMM_Total)
```

#Lactate Dehydrogenase
```{r}
LDHLMM <- lmer(P_CHANGE ~ Time_Point + BFR + Speed + Time_Point*BFR + Time_Point*Speed + (1|Participant), data = LDH_PCHANGE_CL)
summary(LDHLMM)
pairwiseLDH <- emmeans(LDHLMM, pairwise ~ Time_Point + BFR + Speed + Time_Point*BFR + Time_Point*Speed, adjust="holm")
pairwiseLDH
```

```{r}
tab_model(LDHLMM)
```

```{r}
names(main_df2)

```

#Sodium
```{r}
NaLMM <- lmer(cal_Na ~ Time_Point + BFR + Speed + Time_Point*BFR + Time_Point*Speed + (1|Participant), data = main_df2)
summary(NaLMM)
pairwise_NaLMM <- emmeans(NaLMM, pairwise ~ Time_Point + BFR + Speed + Time_Point*BFR + Time_Point*Speed, adjust="holm")
pairwise_NaLMM
```

```{r}
tab_model(NaLMM)
```

#Potassium
```{r}
KLMM <- lmer(cal_K ~ Time_Point + BFR + Speed + Time_Point*BFR + Time_Point*Speed + (1|Participant), data = main_df2)
summary(KLMM)
pairwise_KLMM <- emmeans(KLMM, pairwise ~ Time_Point + BFR + Speed + Time_Point*BFR + Time_Point*Speed, adjust="holm")
pairwise_KLMM
```

```{r}
tab_model(KLMM)
```

#Chlorine

```{r}
ClLMM <- lmer(cal_Cl ~ Time_Point + BFR + Speed + Time_Point*BFR + Time_Point*Speed + (1|Participant), data = main_df2)
summary(ClLMM)
pairwise_ClLMM <- emmeans(ClLMM, pairwise ~ Time_Point + BFR + Speed + Time_Point*BFR + Time_Point*Speed, adjust="holm")
pairwise_ClLMM
```

```{r}
tab_model(ClLMM)
```